Comparing clustering parameters with rOpenScPCA

### 2025-01-14 analyses/hello-clusters/02_compare-clustering-parameters.Rmd
Update analyses/hello-clusters/02_compare-clustering-parameters.Rmd
Co-authored-by: Joshua Shapiro <josh.shapiro@ccdatalab.org>
updates in response to review

`rOpenScPCA::sweep_clusters()` does allow you to specify values for any other parameters.

@@

It can be helpful (although it is not strictly necessary to keep track) to name this list by the varied `nn` parameter:

@@

# Print the plots with patchwork::wrap_plots()
patchwork::wrap_plots(umap_plots)
```

@@

```{r combine cell metrics list}
cell_metrics_df <- dplyr::bind_rows(cell_metric_list)
```

@@

While there does not appear to be a salient difference among silhouette width distributions, it does appear that purity is higher with a higher nearest neighbors parameter.

@@

```{r combine plot stability}
stability_df <- dplyr::bind_rows(stability_list)

@@

```{r sweep two parameters}
# Define vectors of parameters to vary
nn_values <- c(10, 20, 30)
res_values <-c(0.5, 1.0, 1.5)

@@

      # Create a title for the UMAP with both parameters
      umap_title <- glue::glue(
        "nn: {unique(cluster_df$nn)}; res: {unique(cluster_df$resolution)}"
      )

@@

```{r calculate purity two parameters}
purity_df <- cluster_results_list |>
  purrr::map(
    \(cluster_df) {
      rOpenScPCA::calculate_purity(pca_matrix, cluster_df)
    }
  ) |>
  dplyr::bind_rows()
```

We'll add a column `resolution_label` which we'll use to have informative panel titles in the faceted ggplot we make next.

```{r add resolution_label column}
purity_df <- purity_df |>
  dplyr::mutate(resolution_label = glue::glue("Resolution: {resolution}"))
```

```{r visualize purity two parameters}
ggplot(purity_df) +
  aes(x = as.factor(nn), y = purity, fill = as.factor(nn)) +
  geom_boxplot() +
  scale_fill_brewer(palette = "Pastel2") +
  # facet by resolution
  facet_wrap(vars(resolution_label), labeller = label_both) +
  labs(
    x = "Number of nearest neighbors",
    y = "Neighborhood purity"
  ) +
  theme(legend.position = "none")
```

@@

      rOpenScPCA::calculate_stability(
        pca_matrix,
        cluster_df,
        nn = nn,
        resolution = resolution
      )
    }
  ) |>
  dplyr::bind_rows()

stability_df <- stability_df |>
  dplyr::mutate(resolution_label = glue::glue("Resolution: {resolution}"))
```

```{r visualize stability two parameters}
ggplot(stability_df) +
  aes(x = as.factor(nn), y = ari, fill = as.factor(nn)) +
  geom_boxplot() +
  scale_fill_brewer(palette = "Pastel2") +
  # facet by resolution
  facet_wrap(vars(resolution_label)) +
  labs(
    x = "Number of nearest neighbors",
    y = "Adjusted Rand Index"
  ) +
  theme(legend.position = "none")
```

>>>>

`rOpenScPCA::sweep_clusters()` does not allow you to specify values for any other parameters.

@@

It can be helpful (although it is not strictly necessary to keep track) to name this list by the varied `nn` parameter.
In this case, we'll use these names to label plots.

@@

@@

```{r combine cell metrics list}
cell_metrics_df <- purrr::list_rbind(cell_metric_list)
```

@@

While there does not appear to be a salient difference among silhouette width distributions, it does appear that purity is higher with a higher nearest neighbors parameter.
It's worth noting that this trend in purity values is expected: Higher nearest neighbor parameter values lead to fewer clusters, and neighborhood purity tends to be higher when there are fewer clusters.

@@

```{r combine plot stability}
stability_df <- purrr::list_rbind(stability_list)

@@

```{r sweep two parameters}
# Define vectors of parameters to vary
nn_values <- c(10, 20, 30)
res_values <- c(0.5, 1.0, 1.5)

@@

      # Create a title for the UMAP with both parameters
      umap_title <- glue::glue(
        "nn: {cluster_df$nn[1]}; res: {cluster_df$resolution[1]}"
      )

@@

```{r calculate purity two parameters}
purity_df <- cluster_results_list |>
  purrr::map(
    \(cluster_df) {
      rOpenScPCA::calculate_purity(pca_matrix, cluster_df)
    }
  ) |>
  purrr::list_rbind()
```

```{r visualize purity two parameters}
ggplot(purity_df) +
  aes(x = as.factor(nn), y = purity, fill = as.factor(nn)) +
  geom_boxplot() +
  scale_fill_brewer(palette = "Pastel2") +
  # facet by resolution, labeling panels with both the resolution column name and value
  facet_wrap(vars(resolution), labeller = label_both) +
  labs(
    x = "Number of nearest neighbors",
    y = "Neighborhood purity"
  ) +
  theme(legend.position = "none")
```

@@

      rOpenScPCA::calculate_stability(
        pca_matrix,
        cluster_df,
        nn = nn,
        resolution = resolution
      )
    }
  ) |>
  purrr::list_rbind()
```

```{r visualize stability two parameters}
ggplot(stability_df) +
  aes(x = as.factor(nn), y = ari, fill = as.factor(nn)) +
  geom_boxplot() +
  scale_fill_brewer(palette = "Pastel2") +
  facet_wrap(vars(resolution), labeller = label_both) +
  labs(
    x = "Number of nearest neighbors",
    y = "Adjusted Rand Index"
  ) +
  theme(legend.position = "none")
```
